### 2023-10-09 proposal/proposal.tex
upload of proposal latex files

\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage[margin = 1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage[colorlinks=true, citecolor=blue]{hyperref}

\title{Proposal: Predicting Diabetes Based on Various Factors}
\author{Ashley Merritt\\
  Department of Statistics\\
  University of Connecticut
  }

\begin{document}
\maketitle

\paragraph{Introduction}
According to the American Diabetes Association, about 11.3 percent of Americans, or about 37.3 million people, have diabetes. Of those 37.3 million people, about 28.7 million were actually diagnosed with diabetes, while the remaining 8.6 million were left undiagnosed (\citet{CDC2022Diabetes}). Those left undiagnosed are at risk for even more serious illness if left untreated. Diabetes is a disease where your body does not create enough insulin, or use it properly, in order to get glucose into your cells and use it for energy (\cite{NIH2023Whatis}). Therefore, it is very important that your body has a normal glucose level. In this paper, I will be working to determine the pre-existing factors that can be used in order to predict if a person may have diabetes. Establishing the predictors is very important to me as many of my family members have a long history of suffering from this disease. It is important to better predict this disease to save people from unnecessary suffering.

Previously, the prediction of diabetes has been studied using machine learning models. Specifically in Sisodia's paper, it focused on the "Prediction of Diabetes using Classification Algorithms", they worked through the topic by producing results using support vector machine, Naive Bayes classifier, and decision tree classifier (\cite{Sisodia2018Prediction}). In this paper, I will be continuing the investigation with a different data set using logistic regression and various other statistical tests..

\paragraph{Specific Aims}
In this paper the specific research question that I will be focusing on is: how can we use machine learning models in order to identify individuals that may have already or are at risk at developing diabetes? This will be researched using a data set that contains various metrics on one's health that will be described in the next section of the paper.

\paragraph{Data}

In order to study the proposed research question on the topic of diabetes, I searched through many databases in order to find a set that had all of the information I was looking for. After exploration, I found a data set entitled "Healthcare Diabetes Data set" on the website Kaggle. The data set was originally sourced from the National Institute of Diabetes and Digestive and Kidney Diseases. The data encompasses eight different predictors including pregnancies, glucose, blood pressure, skin thickness, insulin, BMI, diabetes pedigree function, and age.

The predictors listed above will be represented in my research as the following variables. 'Pregnancies' will provide the number of times an entry has been pregnant. 'Glucose' will provide the plasma glucose concentration over two hours using the results of an oral glucose tolerance test on a entry. 'BloodPressure' gives the diastolic blood pressure in mm Hg of an entry. 'SkinThickness' will provide the triceps skin fold thickness in mm. 'Insulin' will include a test on two hour serum insulin in mu U/ml. 'BMI' will provide a calculation of weight (kg) divided by height ($m^2$). 'DiabetesPedigreeFunction' will provide a genetic score of diabetes. 'Age' will simply be the entry's age at the time of the collection.

\paragraph{Research Design and Methods}
Various methods will be used to complete the statistical analysis necessary for this project. I will begin by exploring basic descriptive statistics to gain an understanding of my data set. I will then progress to creating visualizations including histograms and box plots to understand relationships and patterns within my data. In order to draw insight on my data, I will perform a logistic regression analysis to determine which predictors are significant in predicting if a patient may have diabetes. Within this analysis I will perform necessary transformations in order to create the best model that I can. From my results I will draw necessary conclusions about my data.

\paragraph{Discussion}

As with any data set, we are always challenged with the fact that the data chosen may not show all possibilities of the entire population as it is only a sample and therefore only a scope. Therefore, we can only come to conclusions based on the data that we have chosen to use. Another challenging part of this task is determining which predictors we are going to use. This will require careful experimentation as well as healthcare knowledge.

As this is a paper dealing with sensitive medical data, we have many limitations in order to be ethical as well as respecting patient privacy. Based on the data set collected we should be able to ensure patient privacy pretty well as names or any very unique identifiers are not listed with each entry. We simply have a ID and age for each entry so we do not have to worry as much for the sake of privacy. In terms of our direct models we may have to disregard some assumptions as they may not fit with real world data. If something unexpected happens we may have to reconsider the data set used and consider finding a new one. This will be done by a case by case scenario and will be investigated thoroughly before making any decisions.

\bibliography{proposal/refs}
\bibliographystyle{apalike}

\end{document}

### 2023-10-31 manuscript/manuscript.tex
Creationg of manuscript text file with initial set up

--- 2023-10-09 proposal/proposal.tex
@@
 \usepackage[colorlinks=true, citecolor=blue]{hyperref}
-
-\title{Proposal: Predicting Diabetes Based on Various Factors}
-\author{Ashley Merritt\\
-  Department of Statistics\\
-  University of Connecticut
-  }
-
-\begin{document}
-\maketitle
-
-\paragraph{Introduction}
-According to the American Diabetes Association, about 11.3 percent of Americans, or about 37.3 million people, have diabetes.
-Of those 37.3 million people, about 28.7 million were actually diagnosed with diabetes, while the remaining 8.6 million were left undiagnosed (\citet{CDC2022Diabetes}).
-Those left undiagnosed are at risk for even more serious illness if left untreated.
-Diabetes is a disease where your body does not create enough insulin, or use it properly, in order to get glucose into your cells and use it for energy (\cite{NIH2023Whatis}).
-Therefore, it is very important that your body has a normal glucose level.
-In this paper, I will be working to determine the pre-existing factors that can be used in order to predict if a person may have diabetes.
-Establishing the predictors is very important to me as many of my family members have a long history of suffering from this disease.
-It is important to better predict this disease to save people from unnecessary suffering.
-
-Previously, the prediction of diabetes has been studied using machine learning models.
-Specifically in Sisodia's paper, it focused on the "Prediction of Diabetes using Classification Algorithms", they worked through the topic by producing results using support vector machine, Naive Bayes classifier, and decision tree classifier (\cite{Sisodia2018Prediction}).
-In this paper, I will be continuing the investigation with a different data set using logistic regression and various other statistical tests..
-
-\paragraph{Specific Aims}
-In this paper the specific research question that I will be focusing on is: how can we use machine learning models in order to identify individuals that may have already or are at risk at developing diabetes?
-This will be researched using a data set that contains various metrics on one's health that will be described in the next section of the paper.
-
-\paragraph{Data}
-
-In order to study the proposed research question on the topic of diabetes, I searched through many databases in order to find a set that had all of the information I was looking for.
-After exploration, I found a data set entitled "Healthcare Diabetes Data set" on the website Kaggle.
-The data set was originally sourced from the National Institute of Diabetes and Digestive and Kidney Diseases.
-The data encompasses eight different predictors including pregnancies, glucose, blood pressure, skin thickness, insulin, BMI, diabetes pedigree function, and age.
-
-The predictors listed above will be represented in my research as the following variables.
-'Pregnancies' will provide the number of times an entry has been pregnant.
-'Glucose' will provide the plasma glucose concentration over two hours using the results of an oral glucose tolerance test on a entry.
-'BloodPressure' gives the diastolic blood pressure in mm Hg of an entry.
-'SkinThickness' will provide the triceps skin fold thickness in mm.
-'Insulin' will include a test on two hour serum insulin in mu U/ml.
-'BMI' will provide a calculation of weight (kg) divided by height ($m^2$).
-'DiabetesPedigreeFunction' will provide a genetic score of diabetes.
-'Age' will simply be the entry's age at the time of the collection.
-
-\paragraph{Research Design and Methods}
-Various methods will be used to complete the statistical analysis necessary for this project.
-I will begin by exploring basic descriptive statistics to gain an understanding of my data set.
-I will then progress to creating visualizations including histograms and box plots to understand relationships and patterns within my data.
-In order to draw insight on my data, I will perform a logistic regression analysis to determine which predictors are significant in predicting if a patient may have diabetes.
-Within this analysis I will perform necessary transformations in order to create the best model that I can.
-From my results I will draw necessary conclusions about my data.
-
-\paragraph{Discussion}
-
-As with any data set, we are always challenged with the fact that the data chosen may not show all possibilities of the entire population as it is only a sample and therefore only a scope.
-Therefore, we can only come to conclusions based on the data that we have chosen to use.
-Another challenging part of this task is determining which predictors we are going to use.
-This will require careful experimentation as well as healthcare knowledge.
-
-As this is a paper dealing with sensitive medical data, we have many limitations in order to be ethical as well as respecting patient privacy.
-Based on the data set collected we should be able to ensure patient privacy pretty well as names or any very unique identifiers are not listed with each entry.
-We simply have a ID and age for each entry so we do not have to worry as much for the sake of privacy.
-In terms of our direct models we may have to disregard some assumptions as they may not fit with real world data.
-If something unexpected happens we may have to reconsider the data set used and consider finding a new one.
-This will be done by a case by case scenario and will be investigated thoroughly before making any decisions.
-
-\bibliography{proposal/refs}
-\bibliographystyle{apalike}
-
-\end{document}

### 2023-10-31 manuscript/manuscript.tex
Creation of title portion of manuscript
Creation of headers in manuscript
Preliminary introduction draft

@@
 \usepackage[colorlinks=true, citecolor=blue]{hyperref}
+
+\title{Predicting Diabetes Based on Various Factors}
+\author{Ashley Merritt\\
+    Department of Statistics\\
+    University of Connecticut
+    }
+
+\begin{document}
+\maketitle
+
+\begin{abstract}
+
+\end{abstract}
+
+\section{Introduction}
+\label{sec:intro}
+According to the American Diabetes Association, about 11.3 percent of Americans, or about 37.3 million people, have diabetes.
+Of those 37.3 million people, about 28.7 million were actually diagnosed with diabetes, while the remaining 8.6 million were left undiagnosed (\citet{CDC2022Diabetes}).
+Those left undiagnosed are at risk for even more serious illness if left untreated.
+Diabetes is a disease where your body does not create enough insulin, or use it properly, in order to get glucose into your cells and use it for energy (\cite{NIH2023Whatis}).
+Therefore, it is very important that your body has a normal glucose level.
+In this paper, I will be working to determine the pre-existing factors that can be used in order to predict if a person may have diabetes.
+Establishing the predictors is very important to me as many of my family members have a long history of suffering from this disease.
+It is important to better predict this disease to save people from unnecessary suffering.
+
+Previously, the prediction of diabetes has been studied using machine learning models.
+Specifically in Sisodia's paper, it focused on the "Prediction of Diabetes using Classification Algorithms", they worked through the topic by producing results using support vector machine, Naive Bayes classifier, and decision tree classifier (\cite{Sisodia2018Prediction}).
+In this paper, I will be continuing the investigation with a different data set using logistic regression and various other statistical tests..
+
+\setion{Data}
+\label{sec:data}
+
+\section{Methods}
+\label{sec:meth}
+
+\section{Results}
+\label{sec:resu}
+
+\section{Discussion}
+\label{sec:disc}
+
+\bibliography{refs}
+\bibliographystyle{mcap}
+
+\end{document}

### 2023-10-31 manuscript/manuscript.tex
Preliminary data explanation and hyperlink

@@
 \label{sec:data}
+In order to study the proposed research question on the topic of diabetes, I searched through many databases in order to find a set that had all of the information I was looking for.
+After exploration, I found a data set entitled "Healthcare Diabetes Data set" on the website Kaggle.
+The data set was originally sourced from the National Institute of Diabetes and Digestive and Kidney Diseases.
+The data encompasses eight different predictors including pregnancies, glucose, blood pressure, skin thickness, insulin, BMI, diabetes pedigree function, and age.
+
+The predictors listed above will be represented in my research as the following variables.
+'Pregnancies' will provide the number of times an entry has been pregnant.
+'Glucose' will provide the plasma glucose concentration over two hours using the results of an oral glucose tolerance test on a entry.
+'BloodPressure' gives the diastolic blood pressure in mm Hg of an entry.
+'SkinThickness' will provide the triceps skin fold thickness in mm.
+'Insulin' will include a test on two hour serum insulin in mu U/ml.
+'BMI' will provide a calculation of weight (kg) divided by height ($m^2$).
+'DiabetesPedigreeFunction' will provide a genetic score of diabetes.
+'Age' will simply be the entry's age at the time of the collection.
+
+Dataset link: \href{https://www.kaggle.com/datasets/nanditapore/healthcare-diabetes}{Healthcare Diabetes}
 

### 2023-10-31 manuscript/manuscript.tex
Minor edit to paper title
Specific aim of the paper written out

@@
 
-\title{Predicting Diabetes Based on Various Factors}
+\title{Predicting Diabetes Based on Various Factors Using logistic Regression}
 \author{Ashley Merritt\\
@@
 In this paper, I will be continuing the investigation with a different data set using logistic regression and various other statistical tests..
+
+In this paper the specific research question that I will be focusing on is: how can we use machine learning models in order to identify individuals that may have already or are at risk at developing diabetes?
+This will be researched using a data set that contains various metrics on one's health that will be described in the next section of the paper.
 

### 2023-11-12 code/Descriptivestatistics.R
Creation of code for table 1

@@ -1,0 +1,11 @@
+desc_stats <- data.frame(
+  Variable = colnames(diabetes),
+  Mean = colMeans(diabetes, na.rm = TRUE),
+  SD = apply(diabetes, 2, sd, na.rm = TRUE),
+  Median = apply(diabetes, 2, median, na.rm = TRUE)
+)
+
+latex_table_xtable <- xtable(desc_stats, caption = "Descriptive Statistics")
+
+print(latex_table_xtable, include.rownames = FALSE)
+

### 2023-11-12 manuscript/manuscript.tex
Small typo in formating
Formation of table 1

@@
 
-\setion{Data}
+\section{Data}
 \label{sec:data}
@@
 
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{lrrr}
+      \hline
+    Variable & Mean & SD & Median \\ 
+      \hline
+    Id & 1384.50 & 799.20 & 1384.50 \\ 
+      Pregnancies & 3.74 & 3.32 & 3.00 \\ 
+      Glucose & 121.10 & 32.04 & 117.00 \\ 
+      BloodPressure & 69.13 & 19.23 & 72.00 \\ 
+      SkinThickness & 20.82 & 16.06 & 23.00 \\ 
+      Insulin & 80.13 & 112.30 & 37.00 \\ 
+      BMI & 32.14 & 8.08 & 32.20 \\ 
+      DiabetesPedigreeFunction & 0.47 & 0.33 & 0.38 \\ 
+      Age & 33.13 & 11.78 & 29.00 \\ 
+      Outcome & 0.34 & 0.48 & 0.00 \\ 
+       \hline
+    \end{tabular}
+    \caption{Descriptive Statistics} 
+    \end{table}
+
 \section{Methods}

### 2023-11-12 manuscript/manuscript.tex
Some changes to formation of Table 1 as well as a reference to it in the text.
minor changes to table 1 and its description in text
Introductory method section paragraph written out.

@@
 'Age' will simply be the entry's age at the time of the collection.
+Following this paragraph, I will provide the link to the dataset that I will be using as well as showing some descriptive statistics in Table~\ref{tab:ds} including the mean, standard deviation, and median.
 
@@
 \begin{table}[ht]
+    \caption{Descriptive Statistics of Healthcare Diabetes Dataset}
+  \label{tab:ds}
 \centering
@@
     \end{tabular}
-    \caption{Descriptive Statistics} 
     \end{table}
@@
 \label{sec:meth}
+Various methods will be used to complete the statistical analysis necessary for this project.
+I will begin by exploring basic descriptive statistics to gain an understanding of my data set.
+I will then progress to creating visualizations including histograms and box plots to understand relationships and patterns within my data.
+In order to draw insight on my data I will perform logistic regression analysis to determine which predictors are significant in predicting if someone has diabetes.
+Within this analysis I will perform necessary transformations in order to create the best model that I can.
+From my results I will draw necessary conclusions about my data.
 

### 2023-11-12 code/Descriptivestatistics.R
Changing of code in order to leave out "ID" and "Outcome" from table 1

@@ -1,8 +1,12 @@
+exclude_vars <- c("Id", "Outcome")
+
+subset_data <- diabetes[, !(names(diabetes) %in% exclude_vars)]
+
 desc_stats <- data.frame(
-  Variable = colnames(diabetes),
-  Mean = colMeans(diabetes, na.rm = TRUE),
-  SD = apply(diabetes, 2, sd, na.rm = TRUE),
-  Median = apply(diabetes, 2, median, na.rm = TRUE)
+  Variable = colnames(subset_data),
+  Mean = colMeans(subset_data, na.rm = TRUE),
+  SD = apply(subset_data, 2, sd, na.rm = TRUE),
+  Median = apply(subset_data, 2, median, na.rm = TRUE)
 )
 
 latex_table_xtable <- xtable(desc_stats, caption = "Descriptive Statistics")

### 2023-11-12 manuscript/manuscript.tex
Minor change to table 1 in order to remove unneeded data

@@
       \hline
-    Id & 1384.50 & 799.20 & 1384.50 \\ 
       Pregnancies & 3.74 & 3.32 & 3.00 \\ 
@@
       Age & 33.13 & 11.78 & 29.00 \\ 
-      Outcome & 0.34 & 0.48 & 0.00 \\ 
        \hline

### 2023-11-13 code/LogisticRegression.R
Code for logistic regression summary table

@@ -1,0 +1,21 @@
+library(xtable)
+
+set.seed(123)
+index <- sample(seq_len(nrow(diabetes)), size = 0.8 * nrow(diabetes))
+
+train_data <- diabetes[index, ]
+test_data <- diabetes[-index, ]
+
+logistic_model <- glm(Outcome ~ ., data = train_data, family = "binomial")
+
+summary_info <- summary(logistic_model)$coefficients[, c("Estimate", "Std. Error", "z value", "Pr(>|z|)")]
+
+predictor_labels <- data.frame(
+  Predictor = names(coef(logistic_model))
+)
+
+summary_df <- cbind(predictor_labels, summary_info)
+
+latex_table <- xtable(summary_df)
+
+print(latex_table, include.rownames = FALSE)

### 2023-11-13 manuscript/manuscript.tex
Formatting issues
Continuation of methods including the creation of table 2

@@
 
+    After exploring the descriptive statistics, I will begin to perform a logistic regression.
+Logistic regression, is used to describe the data and the relationship between one dependent variable against many independent variables.
+In this particular case, we will be looking at the binary response of diabetes with the various independent variables including but not limited to; pregnancies and glucose levels.
+To begin, I will split the data into a training and a test set.
+This will be done by setting a random seed of 30 and using 80 percent of the data set in the training set and the remaining 20 percent as the testing set.
+We will next begin to build the logistic regression model using the training data.
+
+    In Table~\ref{tab:sl} you will see the output of each predictor on predicting the outcome of diabetes.
+This table will provide us with coefficients, the standard errors, the z-scores, and finally the p-value for each predictor in the model.
+Based on the output, we can determine that only skin thickness is not statistically significant in predicting if a patient may have diabetes.
+Therefore, all of the other predictors are statistically significant and should be monitored in predicting diabetes.
+
+\begin{table}[ht]
+    \caption{Summary of Logistic Regression Model}
+    \label{tab:sl}
+\centering
+\begin{tabular}{lrrrr}
+  \hline
+Predictor & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  \hline
+(Intercept) & -8.02 & 0.41 & -19.73 & 0.00 \\ 
+  Pregnancies & 0.11 & 0.02 & 6.04 & 0.00 \\ 
+  Glucose & 0.03 & 0.00 & 16.34 & 0.00 \\ 
+  BloodPressure & -0.01 & 0.00 & -3.69 & 0.00 \\ 
+  SkinThickness & 0.00 & 0.00 & 0.26 & 0.79 \\ 
+  Insulin & -0.00 & 0.00 & -2.74 & 0.01 \\ 
+  BMI & 0.08 & 0.01 & 9.07 & 0.00 \\ 
+  DiabetesPedigreeFunction & 0.76 & 0.17 & 4.38 & 0.00 \\ 
+  Age & 0.02 & 0.01 & 3.03 & 0.00 \\ 
+   \hline
+\end{tabular}
+\end{table}
 \section{Results}

### 2023-11-13 manuscript/manuscript.tex
minor edits
Addition of VIF table and beginning of assumption section

@@
 \end{abstract}
+
+The rest of the paper is organized as follows.
+The data will be presented in Section~\ref{sec:data}.
+The methods are described in Section~\ref{sec:meth}.
+The results are reported in Section~\ref{sec:resu}.
+A discussion concludes in Section~\ref{sec:disc}.
 
@@
 In order to draw insight on my data I will perform logistic regression analysis to determine which predictors are significant in predicting if someone has diabetes.
+In order to perform the logistic regression, I will have to check the necessary assumptions associated with the method.
 Within this analysis I will perform necessary transformations in order to create the best model that I can.
 From my results I will draw necessary conclusions about my data.
+
+    The first assumption that will be checked in that the dependent variable is binary, this true as our outcome variable is in the form of 1 (having diabetes) or 2 (not having diabetes).
+The next assumption that will be checked is if the observations are independent of each other, this is also true for our data as each predictor is calculated or taken independent of one another.
+The third assumption that will be checked is that there is no multicollinearity among the independent variables.
+In order to check this assumption we will calculate the variance inflation factor, VIF, in order to measure the strength of the correlation between predictor variables in our model.
+One example of this is seen in Table~\ref{tab:vif} where all of the values are around 1, indicating that there is moderate correlation between pregnancies and the other predictors but this value is not in a concerning range.
+
+\begin{table}[ht]
+\caption{VIF Values} 
+\label{tab:vif}
+\centering
+\begin{tabular}{rr}
+  \hline
+ & vif(lm(P \~{} G + BP + ST + I + BMI + DPF + A)) \\ 
+  \hline
+Glucose & 1.29 \\ 
+  BloodPressure & 1.18 \\ 
+  SkinThickness & 1.50 \\ 
+  Insulin & 1.44 \\ 
+  BMI & 1.30 \\ 
+  DiabetesPedigreeFunction & 1.06 \\ 
+  Age & 1.18 \\ 
+   \hline
+\end{tabular}
+\end{table}
 

### 2023-11-14 manuscript/manuscript.tex
Start of results section

@@
 
+After completing our methods section, we can now move on to talk about the results of our model.
+Based on our logistic regression model, we now know that all of the predictors in the data set except skin thickness are important in order to predict whether a patient may have diabetes or not.
+With this information medical professionals will be able to ensure proper treatment to patients as they can use these predictors in order to tell if a patient may have diabetes even sooner than typical means.
+The model allows for medical professionals to use the model in order to further their research into diabetes and the various precursors.
+
 \section{Discussion}

### 2023-11-14 manuscript/manuscript.tex
Beginning of discussion section
Continuation of discussion

@@
 
+Currently, this study is limited as we are set within the bounds of this data set.
+In future studies we can include more predictors in order to see if there are any other factors that could be used to predict diabetes.
+As well as having an even larger data set in order to ensure our accuracy.
+Other data sets may suggest that certain predictors are not helpful in predicting diabetes that our model did not pick up on due to the metrics of our data.
+
+As this is a paper dealing with sensitive medical data, we have many limitations in order to be ethical as well as respecting patient privacy.
+Based on the data set collected we should be able to ensure patient privacy pretty well as names or any very unique identifiers are not listed with each entry.
+We simply have a ID and age for each entry so we do not have to worry as much for the sake of privacy.
+In terms of our direct models we may have to disregard some assumptions as they may not fit with real world data.
+If something unexpected happens we may have to reconsider the data set used and consider finding a new one.
+This will be done by a case by case scenario and will be investigated thoroughly before making any decisions.
+
 \bibliography{refs}

### 2023-11-22 manuscript/manuscript.tex
Abstract explanation
Simple formating changes

@@
 \begin{abstract}
-
+This section will be completed once the paper is closer to completion. 
 \end{abstract}
@@
     According to the American Diabetes Association, about 11.3 percent of Americans, or about 37.3 million people, have diabetes.
-Of those 37.3 million people, about 28.7 million were actually diagnosed with diabetes, while the remaining 8.6 million were left undiagnosed (\citet{CDC2022Diabetes}).
-Those left undiagnosed are at risk for even more serious illness if left untreated.
-Diabetes is a disease where your body does not create enough insulin, or use it properly, in order to get glucose into your cells and use it for energy (\cite{NIH2023Whatis}).
-Therefore, it is very important that your body has a normal glucose level.
-In this paper, I will be working to determine the pre-existing factors that can be used in order to predict if a person may have diabetes.
+    Of those 37.3 million people, about 28.7 million were actually diagnosed with diabetes, while the remaining 8.6 million were
+    left undiagnosed (\citet{CDC2022Diabetes}). Those left undiagnosed are at risk for even more serious illness if left untreated. 
+    Diabetes is a disease where your body does not create enough insulin, or use it properly, in order to get glucose into your cells 
+    and use it for energy (\cite{NIH2023Whatis}). Therefore, it is very important that your body has a normal glucose level. In this 
+    paper, I will be working to determine the pre-existing factors that can be used in order to predict if a person may have diabetes. 
     Establishing the predictors is very important to me as many of my family members have a long history of suffering from this disease. 
@@
 
-    Previously, the prediction of diabetes has been studied using machine learning models.
-Specifically in Sisodia's paper, it focused on the "Prediction of Diabetes using Classification Algorithms", they worked through the topic by producing results using support vector machine, Naive Bayes classifier, and decision tree classifier (\cite{Sisodia2018Prediction}).
-In this paper, I will be continuing the investigation with a different data set using logistic regression and various other statistical tests..
+    Previously, the prediction of diabetes has been studied using machine learning models. Specifically in Sisodia's paper, it focused
+    on the "Prediction of Diabetes using Classification Algorithms", they worked through the topic by producing results using support 
+    vector machine, Naive Bayes classifier, and decision tree classifier (\cite{Sisodia2018Prediction}). In this paper, I will be 
+    continuing the investigation with a different data set using logistic regression and various other statistical tests.
 
-    In this paper the specific research question that I will be focusing on is: how can we use machine learning models in order to identify individuals that may have already or are at risk at developing diabetes?
-This will be researched using a data set that contains various metrics on one's health that will be described in the next section of the paper.
+    In this paper the specific research question that I will be focusing on is: how can we use machine learning models in order to identify 
+    individuals that may have already or are at risk at developing diabetes? This will be researched using a data set that contains various 
+    metrics on one's health that will be described in the next section of the paper.
 
@@
 \label{sec:data}
-    In order to study the proposed research question on the topic of diabetes, I searched through many databases in order to find a set that had all of the information I was looking for.
-After exploration, I found a data set entitled "Healthcare Diabetes Data set" on the website Kaggle.
-The data set was originally sourced from the National Institute of Diabetes and Digestive and Kidney Diseases.
-The data encompasses eight different predictors including pregnancies, glucose, blood pressure, skin thickness, insulin, BMI, diabetes pedigree function, and age.
+    In order to study the proposed research question on the topic of diabetes, I searched through many databases in order to find a set 
+    that had all of the information I was looking for. After exploration, I found a data set entitled "Healthcare Diabetes Data set" on 
+    the website Kaggle. The data set was originally sourced from the National Institute of Diabetes and Digestive and Kidney Diseases. The 
+    data encompasses eight different predictors including pregnancies, glucose, blood pressure, skin thickness, insulin, BMI, diabetes 
+    pedigree function, and age.
 
-    The predictors listed above will be represented in my research as the following variables.
-'Pregnancies' will provide the number of times an entry has been pregnant.
-'Glucose' will provide the plasma glucose concentration over two hours using the results of an oral glucose tolerance test on a entry.
-'BloodPressure' gives the diastolic blood pressure in mm Hg of an entry.
-'SkinThickness' will provide the triceps skin fold thickness in mm.
-'Insulin' will include a test on two hour serum insulin in mu U/ml.
-'BMI' will provide a calculation of weight (kg) divided by height ($m^2$).
-'DiabetesPedigreeFunction' will provide a genetic score of diabetes.
-'Age' will simply be the entry's age at the time of the collection.
-Following this paragraph, I will provide the link to the dataset that I will be using as well as showing some descriptive statistics in Table~\ref{tab:ds} including the mean, standard deviation, and median.
+    The predictors listed above will be represented in my research as the following variables. 'Pregnancies' will provide the number of 
+    times an entry has been pregnant. 'Glucose' will provide the plasma glucose concentration over two hours using the results of an oral 
+    glucose tolerance test on a entry. 'BloodPressure' gives the diastolic blood pressure in mm Hg of an entry. 'SkinThickness' will 
+    provide the triceps skin fold thickness in mm. 'Insulin' will include a test on two hour serum insulin in mu U/ml. 'BMI' will provide a 
+    calculation of weight (kg) divided by height ($m^2$). 'DiabetesPedigreeFunction' will provide a genetic score of diabetes. 'Age' will 
+    simply be the entry's age at the time of the collection. Following this paragraph, I will provide the link to the dataset that I will 
+    be using as well as showing some descriptive statistics in Table~\ref{tab:ds} including the mean, standard deviation, and median.
 
@@
 \label{sec:meth}
-    Various methods will be used to complete the statistical analysis necessary for this project.
-I will begin by exploring basic descriptive statistics to gain an understanding of my data set.
-I will then progress to creating visualizations including histograms and box plots to understand relationships and patterns within my data.
-In order to draw insight on my data I will perform logistic regression analysis to determine which predictors are significant in predicting if someone has diabetes.
-In order to perform the logistic regression, I will have to check the necessary assumptions associated with the method.
-Within this analysis I will perform necessary transformations in order to create the best model that I can.
-From my results I will draw necessary conclusions about my data.
+    Various methods will be used to complete the statistical analysis necessary for this project. I will begin by exploring basic 
+    descriptive statistics to gain an understanding of my data set. I will then progress to creating visualizations including histograms 
+    and box plots to understand relationships and patterns within my data. In order to draw insight on my data I will perform logistic 
+    regression analysis to determine which predictors are significant in predicting if someone has diabetes. In order to perform the 
+    logistic regression, I will have to check the necessary assumptions associated with the method. Within this analysis I will perform 
+    necessary transformations in order to create the best model that I can. From my results I will draw necessary conclusions about my data.
 
-    The first assumption that will be checked in that the dependent variable is binary, this true as our outcome variable is in the form of 1 (having diabetes) or 2 (not having diabetes).
-The next assumption that will be checked is if the observations are independent of each other, this is also true for our data as each predictor is calculated or taken independent of one another.
-The third assumption that will be checked is that there is no multicollinearity among the independent variables.
-In order to check this assumption we will calculate the variance inflation factor, VIF, in order to measure the strength of the correlation between predictor variables in our model.
-One example of this is seen in Table~\ref{tab:vif} where all of the values are around 1, indicating that there is moderate correlation between pregnancies and the other predictors but this value is not in a concerning range.
+    The first assumption that will be checked in that the dependent variable is binary, this true as our outcome variable is in the form of 
+    1 (having diabetes) or 2 (not having diabetes). The next assumption that will be checked is if the observations are independent of each 
+    other, this is also true for our data as each predictor is calculated or taken independent of one another. The third assumption that 
+    will be checked is that there is no multicollinearity among the independent variables. In order to check this assumption we will 
+    calculate the variance inflation factor, VIF, in order to measure the strength of the correlation between predictor variables in our 
+    model. One example of this is seen in Table~\ref{tab:vif} where all of the values are around 1, indicating that there is moderate 
+    correlation between pregnancies and the other predictors but this value is not in a concerning range.
 
@@
 
-    After exploring the descriptive statistics, I will begin to perform a logistic regression.
-Logistic regression, is used to describe the data and the relationship between one dependent variable against many independent variables.
-In this particular case, we will be looking at the binary response of diabetes with the various independent variables including but not limited to; pregnancies and glucose levels.
-To begin, I will split the data into a training and a test set.
-This will be done by setting a random seed of 30 and using 80 percent of the data set in the training set and the remaining 20 percent as the testing set.
-We will next begin to build the logistic regression model using the training data.
+    After exploring the descriptive statistics, I will begin to perform a logistic regression. Logistic regression, is used to describe the 
+    data and the relationship between one dependent variable against many independent variables. In this particular case, we will be looking 
+    at the binary response of diabetes with the various independent variables including but not limited to; pregnancies and glucose levels. 
+    To begin, I will split the data into a training and a test set. This will be done by setting a random seed of 30 and using 80 percent 
+    of the data set in the training set and the remaining 20 percent as the testing set. We will next begin to build the logistic 
+    regression model using the training data.
 
-    In Table~\ref{tab:sl} you will see the output of each predictor on predicting the outcome of diabetes.
-This table will provide us with coefficients, the standard errors, the z-scores, and finally the p-value for each predictor in the model.
-Based on the output, we can determine that only skin thickness is not statistically significant in predicting if a patient may have diabetes.
-Therefore, all of the other predictors are statistically significant and should be monitored in predicting diabetes.
+    In Table~\ref{tab:sl} you will see the output of each predictor on predicting the outcome of diabetes. This table will provide us with 
+    coefficients, the standard errors, the z-scores, and finally the p-value for each predictor in the model. Based on the output, we can 
+    determine that only skin thickness is not statistically significant in predicting if a patient may have diabetes. Therefore, all of the 
+    other predictors are statistically significant and should be monitored in predicting diabetes.
 
@@
 
-After completing our methods section, we can now move on to talk about the results of our model.
-Based on our logistic regression model, we now know that all of the predictors in the data set except skin thickness are important in order to predict whether a patient may have diabetes or not.
-With this information medical professionals will be able to ensure proper treatment to patients as they can use these predictors in order to tell if a patient may have diabetes even sooner than typical means.
-The model allows for medical professionals to use the model in order to further their research into diabetes and the various precursors.
+After completing our methods section, we can now move on to talk about the results of our model. Based on our logistic regression model, 
+we now know that all of the predictors in the data set except skin thickness are important in order to predict whether a patient may have 
+diabetes or not. With this information medical professionals will be able to ensure proper treatment to patients as they can use these 
+predictors in order to tell if a patient may have diabetes even sooner than typical means. The model allows for medical professionals to 
+use the model in order to further their research into diabetes and the various precursors.
 
@@
 
-Currently, this study is limited as we are set within the bounds of this data set.
-In future studies we can include more predictors in order to see if there are any other factors that could be used to predict diabetes.
-As well as having an even larger data set in order to ensure our accuracy.
-Other data sets may suggest that certain predictors are not helpful in predicting diabetes that our model did not pick up on due to the metrics of our data.
+Currently, this study is limited as we are set within the bounds of this data set. In future studies we can include more predictors in 
+order to see if there are any other factors that could be used to predict diabetes. As well as having an even larger data set in order to 
+ensure our accuracy. Other data sets may suggest that certain predictors are not helpful in predicting diabetes that our model did not pick 
+up on due to the metrics of our data.
 
-As this is a paper dealing with sensitive medical data, we have many limitations in order to be ethical as well as respecting patient privacy.
-Based on the data set collected we should be able to ensure patient privacy pretty well as names or any very unique identifiers are not listed with each entry.
-We simply have a ID and age for each entry so we do not have to worry as much for the sake of privacy.
-In terms of our direct models we may have to disregard some assumptions as they may not fit with real world data.
-If something unexpected happens we may have to reconsider the data set used and consider finding a new one.
-This will be done by a case by case scenario and will be investigated thoroughly before making any decisions.
+As this is a paper dealing with sensitive medical data, we have many limitations in order to be ethical as well as respecting patient 
+privacy. Based on the data set collected we should be able to ensure patient privacy pretty well as names or any very unique identifiers 
+are not listed with each entry. We simply have a ID and age for each entry so we do not have to worry as much for the sake of privacy. In 
+terms of our direct models we may have to disregard some assumptions as they may not fit with real world data. If something unexpected 
+happens we may have to reconsider the data set used and consider finding a new one. This will be done by a case by case scenario and will 
+be investigated thoroughly before making any decisions.
 

### 2023-12-05 code/Head_of_table.R
Table of head of dataset

@@ -1,0 +1,6 @@
+library(xtable)
+diabetes <- read.csv("/Users/ashleymerritt/Downloads/Healthcare-Diabetes 2.csv")
+head_of_data <- head(diabetes)
+latex_table <- xtable(head_of_data, caption = "Head of the Dataset", label = 
+                        "tab:head_of_data")
+print(latex_table, include.rownames = FALSE)

### 2023-12-05 manuscript/manuscript.tex
Double space pdf
Moving of roadmap
Table of head of dataset
Movement of paragraph in method section
Change to decision tree

@@
 \documentclass[12pt]{article}
+\usepackage{setspace}
+\doublespacing
 
@@
 
-\title{Predicting Diabetes Based on Various Factors Using logistic Regression}
+\title{Predicting Diabetes Based on Various Factors Using Decision Trees}
 \author{Ashley Merritt\\
@@
 \end{abstract}
-
-The rest of the paper is organized as follows.
-The data will be presented in Section~\ref{sec:data}.
-The methods are described in Section~\ref{sec:meth}.
-The results are reported in Section~\ref{sec:resu}.
-A discussion concludes in Section~\ref{sec:disc}.
 
@@
     vector machine, Naive Bayes classifier, and decision tree classifier (\cite{Sisodia2018Prediction}). In this paper, I will be 
-    continuing the investigation with a different data set using logistic regression and various other statistical tests.
+    continuing the investigation with a different data set using decision trees.
 
@@
     metrics on one's health that will be described in the next section of the paper.
+
+    The rest of the paper is organized as follows.
+    The data will be presented in Section~\ref{sec:data}.
+    The methods are described in Section~\ref{sec:meth}.
+    The results are reported in Section~\ref{sec:resu}.
+    A discussion concludes in Section~\ref{sec:disc}.
 
@@
     data encompasses eight different predictors including pregnancies, glucose, blood pressure, skin thickness, insulin, BMI, diabetes 
-    pedigree function, and age.
+    pedigree function, and age. In Table~\ref{tab:head_of_data} I will provide the first couple lines of dataset in order to provide
+    more information on the dataset.
+
+    \begin{table}[ht]
+    \caption{Head of the Dataset}
+    \centering
+    \resizebox{\textwidth}{!}{%
+    \begin{tabular}{rrrrrrrrr}
+      \hline
+    Pregnancies & Glucose & BloodPressure & SkinThickness & Insulin & BMI & DiabetesPedigreeFunction & Age & Outcome \\ 
+      \hline
+      6 & 148 &  72 &  35 &   0 & 33.60 & 0.63 &  50 &   1 \\ 
+        1 &  85 &  66 &  29 &   0 & 26.60 & 0.35 &  31 &   0 \\ 
+        8 & 183 &  64 &   0 &   0 & 23.30 & 0.67 &  32 &   1 \\ 
+        1 &  89 &  66 &  23 &  94 & 28.10 & 0.17 &  21 &   0 \\ 
+        0 & 137 &  40 &  35 & 168 & 43.10 & 2.29 &  33 &   1 \\ 
+        5 & 116 &  74 &   0 &   0 & 25.60 & 0.20 &  30 &   0 \\ 
+       \hline
+    \end{tabular}%
+    } 
+    \label{tab:head_of_data}
+    \end{table}
 
@@
     Various methods will be used to complete the statistical analysis necessary for this project. I will begin by exploring basic 
-    descriptive statistics to gain an understanding of my data set. I will then progress to creating visualizations including histograms 
-    and box plots to understand relationships and patterns within my data. In order to draw insight on my data I will perform logistic 
-    regression analysis to determine which predictors are significant in predicting if someone has diabetes. In order to perform the 
-    logistic regression, I will have to check the necessary assumptions associated with the method. Within this analysis I will perform 
-    necessary transformations in order to create the best model that I can. From my results I will draw necessary conclusions about my data.
+    descriptive statistics to gain an understanding of my data set. I will then begin to create my model for my decision tree by
+    splitting the data into a test and training set. The training set will encompass 80 percent of the data whereas the test set will
+    only have 20 percent. After creating these two sets we will now be ready to build the model, this will be done by using the rpart
+    function in R, where method will be 'class' as this is a binary model. We will then see in Figure~\ref{tab:structure} the structure
+    of the decision tree created. Looking at the leftmost path in the decision tree we can see that at the top the outcome was 0 meaning 
+    that the patient did not have diabetes 35 percent of the time. Moving down to the next node, we can see that 68 percent of patients with
+    glucose level less than 128, did not have diabetes 19 percent of the time. To finish this path, 61 percent of the time if the patient was
+    older than 29 years old, then 8 percent of the time the patient had diabetes. The rest of the structure is similar but with different 
+    values and predictors.
 
-    The first assumption that will be checked in that the dependent variable is binary, this true as our outcome variable is in the form of 
-    1 (having diabetes) or 2 (not having diabetes). The next assumption that will be checked is if the observations are independent of each 
-    other, this is also true for our data as each predictor is calculated or taken independent of one another. The third assumption that 
-    will be checked is that there is no multicollinearity among the independent variables. In order to check this assumption we will 
-    calculate the variance inflation factor, VIF, in order to measure the strength of the correlation between predictor variables in our 
-    model. One example of this is seen in Table~\ref{tab:vif} where all of the values are around 1, indicating that there is moderate 
-    correlation between pregnancies and the other predictors but this value is not in a concerning range.
+\begin{figure}[tbp]
+  \centering
+  \caption{Decision Tree Structure}
+  \includegraphics[width=\textwidth]{Decision Tree Structure.pdf}
+  \label{fig:structure}
+  \end{figure}
 
-\begin{table}[ht]
-\caption{VIF Values} 
-\label{tab:vif}
-\centering
-\begin{tabular}{rr}
-  \hline
- & vif(lm(P \~{} G + BP + ST + I + BMI + DPF + A)) \\ 
-  \hline
-Glucose & 1.29 \\ 
-  BloodPressure & 1.18 \\ 
-  SkinThickness & 1.50 \\ 
-  Insulin & 1.44 \\ 
-  BMI & 1.30 \\ 
-  DiabetesPedigreeFunction & 1.06 \\ 
-  Age & 1.18 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-    After exploring the descriptive statistics, I will begin to perform a logistic regression. Logistic regression, is used to describe the 
-    data and the relationship between one dependent variable against many independent variables. In this particular case, we will be looking 
-    at the binary response of diabetes with the various independent variables including but not limited to; pregnancies and glucose levels. 
-    To begin, I will split the data into a training and a test set. This will be done by setting a random seed of 30 and using 80 percent 
-    of the data set in the training set and the remaining 20 percent as the testing set. We will next begin to build the logistic 
-    regression model using the training data.
-
-    In Table~\ref{tab:sl} you will see the output of each predictor on predicting the outcome of diabetes. This table will provide us with 
-    coefficients, the standard errors, the z-scores, and finally the p-value for each predictor in the model. Based on the output, we can 
-    determine that only skin thickness is not statistically significant in predicting if a patient may have diabetes. Therefore, all of the 
-    other predictors are statistically significant and should be monitored in predicting diabetes.
-
-\begin{table}[ht]
-    \caption{Summary of Logistic Regression Model}
-    \label{tab:sl}
-\centering
-\begin{tabular}{lrrrr}
-  \hline
-Predictor & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
-  \hline
-(Intercept) & -8.02 & 0.41 & -19.73 & 0.00 \\ 
-  Pregnancies & 0.11 & 0.02 & 6.04 & 0.00 \\ 
-  Glucose & 0.03 & 0.00 & 16.34 & 0.00 \\ 
-  BloodPressure & -0.01 & 0.00 & -3.69 & 0.00 \\ 
-  SkinThickness & 0.00 & 0.00 & 0.26 & 0.79 \\ 
-  Insulin & -0.00 & 0.00 & -2.74 & 0.01 \\ 
-  BMI & 0.08 & 0.01 & 9.07 & 0.00 \\ 
-  DiabetesPedigreeFunction & 0.76 & 0.17 & 4.38 & 0.00 \\ 
-  Age & 0.02 & 0.01 & 3.03 & 0.00 \\ 
-   \hline
-\end{tabular}
-\end{table}
 \section{Results}

### 2023-12-05 code/Decision Tree Structure Code.R

@@ -1,0 +1,13 @@
+library(tidyverse)
+library(rpart)
+library(rpart.plot)
+
+set.seed(30)
+train_indices <- sample(seq_len(nrow(diabetes)), size = 0.8 * nrow(diabetes))
+train_data <- diabetes[train_indices, ]
+test_data <- diabetes[-train_indices, ]
+
+tree_model <- rpart(Outcome ~ Pregnancies + Glucose + BloodPressure + SkinThickness 
+             + Insulin + BMI + DiabetesPedigreeFunction + Age, 
+             data = train_data, method = "class")
+rpart.plot(tree_model, box.palette = "auto", main = "Decision Tree Structure")

### 2023-12-05 code/LogisticRegression.R

@@ -1,21 +1,0 @@
-library(xtable)
-
-set.seed(123)
-index <- sample(seq_len(nrow(diabetes)), size = 0.8 * nrow(diabetes))
-
-train_data <- diabetes[index, ]
-test_data <- diabetes[-index, ]
-
-logistic_model <- glm(Outcome ~ ., data = train_data, family = "binomial")
-
-summary_info <- summary(logistic_model)$coefficients[, c("Estimate", "Std. Error", "z value", "Pr(>|z|)")]
-
-predictor_labels <- data.frame(
-  Predictor = names(coef(logistic_model))
-)
-
-summary_df <- cbind(predictor_labels, summary_info)
-
-latex_table <- xtable(summary_df)
-
-print(latex_table, include.rownames = FALSE)

### 2023-12-05 code/VIFCorrelation.R

@@ -1,11 +1,0 @@
-library(car)
-library(xtable)
-diabetes$Id <- NULL
-
-cor_matrix <- cor(diabetes[, c('Pregnancies', 'Glucose', 'BloodPressure', 'SkinThickness', 'Insulin', 'BMI', 'DiabetesPedigreeFunction', 'Age')])
-
-print(xtable(cor_matrix, caption = "Correlation Matrix", label = "tab:correlation"), include.rownames = TRUE)
-
-vif_data <- as.data.frame(vif(lm(Pregnancies ~ Glucose + BloodPressure + SkinThickness + Insulin + BMI + DiabetesPedigreeFunction + Age, data = diabetes)))
-
-print(xtable(vif_data, caption = "VIF Values", label = "tab:vif"), include.rownames = TRUE)

### 2023-12-05 code/Confusion Matrix.R
Addition of Confusion Matrix

@@ -1,0 +1,17 @@
+library(rpart)
+library(rpart.plot)
+library(gplots)
+
+predictions <- predict(tree_model, test_data, type = "class")
+conf_matrix <- table(Actual = test_data$Outcome, Predicted = predictions)
+accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
+
+conf_matrix <- table(predictions, test_data$Outcome)
+accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
+
+cat("\\begin{table}[h]\n")
+cat("\\centering\n")
+print(xtable(conf_matrix, caption = "Confusion Matrix", label = "tab:conf_matrix", add.to.row = list(pos = c(-1, 0, nrow(conf_matrix) + 1), command = c("\\toprule ", "\\midrule ", "\\bottomrule ")), digits = 0), include.rownames = TRUE, booktabs = TRUE)
+cat("\\end{table}\n")
+
+

### 2023-12-05 manuscript/manuscript.tex
formatting change
Addition of Confusion Matrix
Title edit

@@
 
-\title{Predicting Diabetes Based on Various Factors Using Decision Trees}
+\title{Predicting Diabetes Based on Various Factors Using Machine Learning Models}
 \author{Ashley Merritt\\
@@
     only have 20 percent. After creating these two sets we will now be ready to build the model, this will be done by using the rpart
-    function in R, where method will be 'class' as this is a binary model. We will then see in Figure~\ref{tab:structure} the structure
+    function in R, where method will be 'class' as this is a binary model. We will then see in Figure~\ref{fig:structure} the structure
     of the decision tree created. Looking at the leftmost path in the decision tree we can see that at the top the outcome was 0 meaning 
@@
   \end{figure}
+
+\begin{table}[ht]
+  \centering
+  \caption{Confusion Matrix} 
+  \begin{tabular}{rrr}
+    \toprule
+    & 0 & 1 \\ 
+    \midrule
+  0 & 318 & 47 \\ 
+    1 & 53 & 136 \\ 
+    \bottomrule
+  \end{tabular}
+  \label{tab:conf_matrix}
+  \end{table}
 

### 2023-12-05 code/Pruned_tree.R
Pruned Tree

@@ -1,0 +1,5 @@
+library(rpart)
+library(rpart.plot)
+pruned_tree <- prune(tree_model, cp =0.01)
+
+rpart.plot(pruned_tree, extra = 4, cex = 0.45)

### 2023-12-05 manuscript/manuscript.tex
Pruned Tree
Decision Tree continued

@@
 \label{sec:meth}
-    Various methods will be used to complete the statistical analysis necessary for this project. I will begin by exploring basic 
-    descriptive statistics to gain an understanding of my data set. I will then begin to create my model for my decision tree by
+    Various methods will be used to complete the statistical analysis necessary for this paper. I will begin by exploring basic 
+    descriptive statistics to gain an understanding of my data set. We will move on to model testing, where we will perform both 
+    decision tree analysis and random forest in order to determine what predictors prove to be necessary in predicting if a patient
+    will have the outcome 1 of having diabetes. Decision trees were choosen to represent the data as they are generally pretty easy
+    to understand.
+
+    To begin, I will provide an explanation of what a decision tree is and why it is useful in this case. A decision tree is considered
+    to be example of supervised learning used for regression and classification. Through our analysis we will present trees, these trees
+    begin at the top with a decision root node, which is the point where you make a decision. Following down there are subsequent sub trees
+    that have other decision nodes. Finally at the bottom of each tree you reach a leaf node, where it is the output of the decision and 
+    there are no further branches off of this.
+
+    I will then begin to create my model for my decision tree by
     splitting the data into a test and training set. The training set will encompass 80 percent of the data whereas the test set will
@@
 
+\begin{figure}[tbp]
+  \centering
+  \caption{Pruned Decision Tree Structure}
+  \includegraphics[width=\textwidth]{Pruned_tree.pdf}
+  \label{fig:prstructure}
+  \end{figure}
+
 \section{Results}

### 2023-12-06 code/Histograms.R
Addition of Histograms, creation of data folder as well as upload

@@ -1,0 +1,16 @@
+library(tidymodels)
+library(tidyr)
+
+diabetes_data_long <- diabetes %>%
+  pivot_longer(cols = everything(),
+               names_to = "variable",
+               values_to = "value")
+diabetes_histograms <- ggplot(diabetes_data_long, aes(x = value)) +
+  geom_histogram(bins = 30, color = "blue", fill = "lightblue") +
+  facet_wrap(~variable, scales = "free", ncol = 4) +
+  labs(title = "Histograms of Variables in Diabetes Dataset",
+       x = "Value",
+       y = "Frequency") + 
+  theme_minimal()
+
+print(diabetes_histograms)

### 2023-12-06 code/Decision Tree Model Performance.R
Decision Tree Model Performance

@@ -1,0 +1,21 @@
+library(tidymodels)
+library(knitr)
+
+tree_spec <- decision_tree() %>%
+  set_engine("rpart") %>%
+  set_mode("regression")
+
+tree_fit <- tree_spec %>%
+  fit(Outcome ~ ., data = train_data)
+
+predictions <- tree_fit %>%
+  predict(test_data) %>%
+  pull(.pred)
+
+metrics <- metric_set(rmse, rsq)
+model_performance <- test_data %>%
+  mutate(predictions = predictions) %>%
+  metrics(truth = test_data$Outcome, estimate = predictions)
+
+latex_table1 <- kable(model_performance, format = "latex", booktabs = TRUE)
+cat(latex_table1)

### 2023-12-06 code/Decision Tree Structure Code.R
Decision tree structure code change

@@ -3,11 +3,14 @@ library(rpart)
 library(rpart.plot)
 
 set.seed(30)
-train_indices <- sample(seq_len(nrow(diabetes)), size = 0.8 * nrow(diabetes))
+train_indices <- sample(1:(nrow(diabetes)), 0.8 * nrow(diabetes))
 train_data <- diabetes[train_indices, ]
 test_data <- diabetes[-train_indices, ]
 
-tree_model <- rpart(Outcome ~ Pregnancies + Glucose + BloodPressure + SkinThickness 
-             + Insulin + BMI + DiabetesPedigreeFunction + Age, 
-             data = train_data, method = "class")
+tree_model <- rpart(Outcome ~ ., data = train_data, method = "class", control = rpart.control(cp = 0.01))
 rpart.plot(tree_model, box.palette = "auto", main = "Decision Tree Structure")
+
+pdf("decision_tree_plot.pdf")
+rpart.plot(tree_model, box.palette = "auto", main = "Decision Tree Structure")
+dev.off()
+

### 2023-12-06 code/Confusion Matrix.R
Decision Tree Results Table, elimination of old model

@@ -1,17 +1,0 @@
-library(rpart)
-library(rpart.plot)
-library(gplots)
-
-predictions <- predict(tree_model, test_data, type = "class")
-conf_matrix <- table(Actual = test_data$Outcome, Predicted = predictions)
-accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
-
-conf_matrix <- table(predictions, test_data$Outcome)
-accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
-
-cat("\\begin{table}[h]\n")
-cat("\\centering\n")
-print(xtable(conf_matrix, caption = "Confusion Matrix", label = "tab:conf_matrix", add.to.row = list(pos = c(-1, 0, nrow(conf_matrix) + 1), command = c("\\toprule ", "\\midrule ", "\\bottomrule ")), digits = 0), include.rownames = TRUE, booktabs = TRUE)
-cat("\\end{table}\n")
-
-

### 2023-12-06 code/Decision Tree Model Performance.R

@@ -1,21 +1,0 @@
-library(tidymodels)
-library(knitr)
-
-tree_spec <- decision_tree() %>%
-  set_engine("rpart") %>%
-  set_mode("regression")
-
-tree_fit <- tree_spec %>%
-  fit(Outcome ~ ., data = train_data)
-
-predictions <- tree_fit %>%
-  predict(test_data) %>%
-  pull(.pred)
-
-metrics <- metric_set(rmse, rsq)
-model_performance <- test_data %>%
-  mutate(predictions = predictions) %>%
-  metrics(truth = test_data$Outcome, estimate = predictions)
-
-latex_table1 <- kable(model_performance, format = "latex", booktabs = TRUE)
-cat(latex_table1)

### 2023-12-06 code/Decision Tree Results.R

@@ -1,0 +1,23 @@
+library(knitr)
+
+predictions <- predict(tree_model, test_data, type = "class")
+conf_matrix <- table(test_data$Outcome, predictions)
+accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
+precision <- conf_matrix[2, 2] / sum(conf_matrix[, 2])
+recall <- conf_matrix[2, 2] / sum(conf_matrix[2, ])
+f1_score <- 2 * (precision * recall) / (precision + recall)
+
+metrics_df <- data.frame( Metric = c("Accuracy", "Precision", "Recall",
+                                     "F1 Score"), Value = c(round(accuracy,
+                                                                  2),
+                                                            round(precision, 2),
+                                                            round(recall, 2),
+                                                            round(f1_score, 2))
+                          )
+latex_code <- knitr::kable(metrics_df, format = "latex", col.names = c("Metric", "Value"), booktabs = TRUE, escape = FALSE)
+cat(latex_code)
+
+
+
+
+

### 2023-12-06 code/Pruned_tree.R

@@ -1,5 +1,0 @@
-library(rpart)
-library(rpart.plot)
-pruned_tree <- prune(tree_model, cp =0.01)
-
-rpart.plot(pruned_tree, extra = 4, cex = 0.45)

### 2023-12-06 code/Confusion Matrix Heatmap.R
Confusion matrix heatmap

@@ -1,0 +1,14 @@
+predictions <- predict(tree_model, newdata = test_data, type = "class")
+
+test_data$Outcome <- factor(test_data$Outcome, levels = levels(predictions))
+
+conf_matrix <- table(Actual = test_data$Outcome, Predicted = predictions)
+
+library(ggplot2)
+ggplot(as.data.frame.table(conf_matrix), aes(x = Actual, y = Predicted, fill = Freq)) +
+  geom_tile() +
+  scale_fill_gradient(low = "white", high = "blue") +
+  geom_text(aes(label = Freq), vjust = 1) +
+  labs(title = "Confusion Matrix Heatmap", x = "Actual Outcome", y = "Predicted Outcome")
+
+

### 2023-12-13 manuscript/manuscript.tex
Addition of Histograms, creation of data folder as well as upload
Decision Tree Model Performance
Decision tree structure code change
Decision Tree Results Table, elimination of old model
Confusion matrix heatmap
Changes and additions to introduction

@@
     Diabetes is a disease where your body does not create enough insulin, or use it properly, in order to get glucose into your cells 
-    and use it for energy (\cite{NIH2023Whatis}). Therefore, it is very important that your body has a normal glucose level. In this 
+    and use it for energy (\cite{NIH2023Whatis}).
+
+    There are two types of diabetes that the majority of the patients have, type 1 or 2. The bodies of people that have type 1 diabetes
+    completely do not make insulin, so they must take a daily insulin shot. Whereas in type 2 diabetes, these patients produce insulin but 
+    their bodies do not use it properly (\cite{JDC2023Difference}).
+
+    Therefore, it is very important that your body has a normal glucose level. In this 
     paper, I will be working to determine the pre-existing factors that can be used in order to predict if a person may have diabetes. 
@@
 
+  \begin{figure}[tbp]
+    \centering
+    \caption{Each variable represented as a histogram}
+      \includegraphics[width=\textwidth]{Histograms.pdf}
+      \label{fig:histogram}
+      \end{figure}
+
 \section{Methods}
@@
     \centering
-  \caption{Confusion Matrix} 
+    \caption{Decision Tree Results}
+    \label{tab:dtr} 
     \begin{tabular}{rrr}
       \toprule
-    & 0 & 1 \\ 
+      Metric & Value\\
       \midrule
-  0 & 318 & 47 \\ 
-    1 & 53 & 136 \\ 
+      Accuracy & 0.82\\
+      Precision & 0.72\\
+      Recall & 0.74\\
+      F1 Score & 0.73\\
       \bottomrule
       \end{tabular}
-  \label{tab:conf_matrix}
     \end{table}
@@
       \centering
-  \caption{Pruned Decision Tree Structure}
-  \includegraphics[width=\textwidth]{Pruned_tree.pdf}
-  \label{fig:prstructure}
+      \caption{Confusion Matrix Heatmap}
+      \includegraphics[width=\textwidth]{Confusion Matrix Heatmap}
+      \label{fig:Heatmap}
       \end{figure}

### 2023-12-13 manuscript/manuscript.tex
Intro formatting
Data section updates

@@
     completely do not make insulin, so they must take a daily insulin shot. Whereas in type 2 diabetes, these patients produce insulin but 
-    their bodies do not use it properly (\cite{JDC2023Difference}).
-
-    Therefore, it is very important that your body has a normal glucose level. In this 
-    paper, I will be working to determine the pre-existing factors that can be used in order to predict if a person may have diabetes. 
+    their bodies do not use it properly (\cite{JDC2023Difference}). Therefore, it is very important that your body has a normal glucose level.
+    In this paper, I will be working to determine the pre-existing factors that can be used in order to predict if a person may have diabetes. 
     Establishing the predictors is very important to me as many of my family members have a long history of suffering from this disease. 
@@
     \end{table}
+  Another valuable piece of information for understanding our patients is seeing the range of the dataset. In Figure~\ref{fig:histogram}, we
+  are able to see the spread of each predictor in our dataset. We can see that the majority of our patients are in the 20 to 30 year old range. 
+  The bloodpressure predictor provides us with a somewhat normal distribution with the average being around 70 beats per minute. The BMI
+  distribution is somewhat normal as well, with a slight right skew. Whereas the diabetes pedigree function is very skewed to the right. Glucose
+  levels appear to be normal. Insulin levels appear to be very skewed to the right with the majority of values being 0. The outcome distribution
+  is exactly as expected with only values of 0 or 1, from here on values of 0 will denote no detection of diabetes and 1 will detect diabetes. The
+  number of pregnancies predictor also supplied a right skew but we know this to be valid as most people are not having 4 or more children. The 
+  final predictor graphed was the skin thickness, this predictor displayed a right skew.
 

### 2023-12-13 manuscript/manuscript.tex
Methods sections continued

@@
 \label{sec:meth}
-    Various methods will be used to complete the statistical analysis necessary for this paper. I will begin by exploring basic 
-    descriptive statistics to gain an understanding of my data set. We will move on to model testing, where we will perform both 
-    decision tree analysis and random forest in order to determine what predictors prove to be necessary in predicting if a patient
-    will have the outcome 1 of having diabetes. Decision trees were choosen to represent the data as they are generally pretty easy
-    to understand.
+    Various methods will be used to complete the analysis necessary for this paper. I have begun by exploring basic descriptive statistics in
+    the previous section in order to gain an understanding of my data set. Next, we will move on to model testing, where we will perform both 
+    decision tree analysis in order to determine what predictors prove to be necessary in predicting if a patient
+    will have the outcome 1 of having diabetes.
 
@@
 
-    I will then begin to create my model for my decision tree by
-    splitting the data into a test and training set. The training set will encompass 80 percent of the data whereas the test set will
-    only have 20 percent. After creating these two sets we will now be ready to build the model, this will be done by using the rpart
-    function in R, where method will be 'class' as this is a binary model. We will then see in Figure~\ref{fig:structure} the structure
-    of the decision tree created. Looking at the leftmost path in the decision tree we can see that at the top the outcome was 0 meaning 
-    that the patient did not have diabetes 35 percent of the time. Moving down to the next node, we can see that 68 percent of patients with
-    glucose level less than 128, did not have diabetes 19 percent of the time. To finish this path, 61 percent of the time if the patient was
-    older than 29 years old, then 8 percent of the time the patient had diabetes. The rest of the structure is similar but with different 
-    values and predictors.
+    I will then begin to create my model for my decision tree by splitting the data into a test and training set. The training set will
+    encompass 80 percent of the data whereas the test set will only have 20 percent. This method of splitting will help to prevent any biases that may 
+    be attributed from imbalances in the distribution of critical features. After creating these two sets we will now be ready
+    to build the model, this will be done by using the rpart function in R, where method will be 'class' as this is a binary model. We will
+    then see in Figure~\ref{fig:structure} the structure of the decision tree created. Looking at the leftmost path in the decision tree we 
+    can see that at the top the outcome was 0 meaning that the patient did not have diabetes 35 percent of the time. Moving down to the next node, 
+    we can see that 68 percent of patients with glucose level less than 128, did not have diabetes 19 percent of the time. To finish this path,
+    61 percent of the time if the patient was older than 29 years old, then 8 percent of the time the patient had diabetes. The rest of the structure
+    is similar but with different values and predictors.
 
@@
 
+  \begin{figure}[tbp]
+    \centering
+    \caption{Variance Importance Plot}
+    \includegraphics[width=\textwidth]{Variance Importance}
+    \label{fig:importance}
+  \end{figure}
+
 \section{Results}

### 2023-12-13 code/Variance Importance.R

@@ -1,0 +1,4 @@
+library(vip)
+
+var_importance <- vip::vip(tree_model)
+print(var_importance)

### 2023-12-13 manuscript/manuscript.tex
grammar edits
Results section edits

@@
       \caption{Confusion Matrix Heatmap}
-      \includegraphics[width=\textwidth]{Confusion Matrix Heatmap}
+      \includegraphics[width=\textwidth]{Confusion Matrix Heatmap.pdf}
       \label{fig:Heatmap}
@@
     \caption{Variance Importance Plot}
-    \includegraphics[width=\textwidth]{Variance Importance}
+    \includegraphics[width=\textwidth]{Variance Important.pdf}
     \label{fig:importance}
@@
 
-After completing our methods section, we can now move on to talk about the results of our model. Based on our logistic regression model, 
-we now know that all of the predictors in the data set except skin thickness are important in order to predict whether a patient may have 
-diabetes or not. With this information medical professionals will be able to ensure proper treatment to patients as they can use these 
-predictors in order to tell if a patient may have diabetes even sooner than typical means. The model allows for medical professionals to 
-use the model in order to further their research into diabetes and the various precursors.
+After completing our methods section, we can now move on to talk about the results of our model. Based on Figure~ref{fig:importance}, we can 
+see that the top three predictors are glucose, BMI, and age. The result of glucose being of the greatest importance makes complete sense with 
+medical knowledge of diabetes as a patient that has high amounts of glucose in their bloodstream could have type 2 diabetes. This is known due
+to glucose not being absorbed by insulin as these patients do not use insulin properly. We can see that almost all of the outcomes are explained
+by glucose. The next predictor that is successful in predicting a large amount of the data is the patients BMI. This alligns on a medical standpoint
+as patients who have larger BMI values are more susceptable to getting type 2 diabetes as they are considered to be obese. Based on our results, BMI
+was able to predict about half of the outcomes. The last predictor that proved to be valuable in predicting if someone has diabetes is age. This result
+is somewhat medically backed but not entirely a result that we expected. Often times people that are older have diabetes due to numerous factors but it is 
+not as common for younger folks.
 

### 2023-12-13 manuscript/manuscript.tex
Abstract formation

@@
 \begin{abstract}
-This section will be completed once the paper is closer to completion. 
+Each year, many patients are given the diagnosis of having diabetes and are left wondering what they could have done to prevent it. This
+paper uses decision tree modeling to investigate which predictors of the Healthcare Diabetes dataset from kaggle are important in predicting 
+an outcome of diabetes. In this study, after various exploratory data analysis, the data will be split into a training and testing set
+that will be used to build a decision tree. The performance was evaluated using standard classification metrics, including accuracy, precision, 
+recall, and F1 score. Our results provided that glucose level emerged as a robust predictor of diabetes. This study contributed to the growing body of
+literature on diabetes prediction. These findings allow for early intervention measures, that can be used to provide a foundation for future research in diabetes
+risk assessment.
 \end{abstract}

### 2023-12-13 manuscript/manuscript.tex
Discussion advances

@@
 
+This study be expanded further by using other machine learning techniques to study the data, this can be used to cross-validate our results to ensure
+that we are providing the most accurate information. Other machine learning techniques could include random forest and logistic regression, to give another
+understanding of the data.
+
 As this is a paper dealing with sensitive medical data, we have many limitations in order to be ethical as well as respecting patient 
 privacy. Based on the data set collected we should be able to ensure patient privacy pretty well as names or any very unique identifiers 
-are not listed with each entry. We simply have a ID and age for each entry so we do not have to worry as much for the sake of privacy. In 
-terms of our direct models we may have to disregard some assumptions as they may not fit with real world data. If something unexpected 
+are not listed with each entry. We simply have age for each entry so we do not have to worry as much for the sake of privacy as this cannot be traced
+back to a particular person. In terms of our direct models we may have to disregard some assumptions as they may not fit with real world data. If something unexpected 
 happens we may have to reconsider the data set used and consider finding a new one. This will be done by a case by case scenario and will 

### 2023-12-13 manuscript/manuscript.tex
minor formatting
Continue Results

@@
     \end{table}
+
   Another valuable piece of information for understanding our patients is seeing the range of the dataset. In Figure~\ref{fig:histogram}, we
@@
 
+Looking at Table~ref{tab:dtr}, we can see that the accuracy of our model is 0.82, this is pretty high so it will be satisfactory. This tells us that we 
+classify correctly 82 percent of the time out of the total instances. The next metric we will look at is precision, we were able to get to 72 percent.
+This is the ratio of correctly predicted positive observations to the total predicted positives. Looking at recall, we were able to correctly identify about
+74 percent of the actual positive instances. This is pretty good and will be considered satisfactory. The final metric that we have is the F1 score, this is the
+harmonic mean of precision and recall, which we were able to get a value of 0.73, this indicates a balanced performance.
+
 \section{Discussion}

### 2023-12-14 manuscript/manuscript.tex
Title edit
Minor edits of intro and data

@@
 \usepackage{setspace}
-\doublespacing
+\singlespacing
 
@@
 
-\title{Predicting Diabetes Based on Various Factors Using Machine Learning Models}
+\title{Predicting Diabetes Based on Various Factors Using Decision Tree Analysis}
 \author{Ashley Merritt\\
@@
 
-    There are two types of diabetes that the majority of the patients have, type 1 or 2. The bodies of people that have type 1 diabetes
-    completely do not make insulin, so they must take a daily insulin shot. Whereas in type 2 diabetes, these patients produce insulin but 
-    their bodies do not use it properly (\cite{JDC2023Difference}). Therefore, it is very important that your body has a normal glucose level.
-    In this paper, I will be working to determine the pre-existing factors that can be used in order to predict if a person may have diabetes. 
-    Establishing the predictors is very important to me as many of my family members have a long history of suffering from this disease. 
-    It is important to better predict this disease to save people from unnecessary suffering.
+    There are two types of diabetes that the majority of patients have, type 1 or 2. The bodies of people that have type 1 diabetes
+    completely do not make insulin, so they must take a daily insulin shot. Whereas individuals who have type 2 diabetes, produce insulin but 
+    their bodies do not use it properly so they are treated either by exercise, taking on a diet or in some cases use of various medicines and/or insulin (\cite{JDC2023Difference}). 
+    Therefore, it is very important that your body has a normal glucose level. In this paper, I will be working to determine the pre-existing
+    factors that can be used in order to predict if a person may have diabetes. Establishing the predictors is very important to me as many of
+    my family members have a long history of suffering from this disease. Therefore, this information could directly allow myself and my family members be
+    proactive in terms of prevention and care. It is important to better predict this disease to save people from unnecessary suffering.
 
-    Previously, the prediction of diabetes has been studied using machine learning models. Specifically in Sisodia's paper, it focused
+    Previously, the prediction of diabetes has been studied using other machine learning models. Specifically in Sisodia's paper, it focused
     on the "Prediction of Diabetes using Classification Algorithms", they worked through the topic by producing results using support 
-    vector machine, Naive Bayes classifier, and decision tree classifier (\cite{Sisodia2018Prediction}). In this paper, I will be 
-    continuing the investigation with a different data set using decision trees.
+    vector machine and Naive Bayes classifier. In Sisodia's paper they were able to get to an accuracy level of 76.3 percent, which we hope to exceed using
+    a different machine learning model (\cite{Sisodia2018Prediction}). In this paper, I will be 
+    continuing the investigation with a different data set using decision trees to better understand the predictors that help to determine if a patient
+    may have diabetes.
 
-    In this paper the specific research question that I will be focusing on is: how can we use machine learning models in order to identify 
+    In this paper the specific research question that I will be focusing on is: how can we use machine learning models, specifically decision trees, in order to identify 
     individuals that may have already or are at risk at developing diabetes? This will be researched using a data set that contains various 
@@
     In order to study the proposed research question on the topic of diabetes, I searched through many databases in order to find a set 
-    that had all of the information I was looking for. After exploration, I found a data set entitled "Healthcare Diabetes Data set" on 
+    that had all of the information I was looking for. After exploration, I was able to find a data set entitled "Healthcare Diabetes Data set" on 
     the website Kaggle. The data set was originally sourced from the National Institute of Diabetes and Digestive and Kidney Diseases. The 
@@
     pedigree function, and age. In Table~\ref{tab:head_of_data} I will provide the first couple lines of dataset in order to provide
-    more information on the dataset.
+    more information.
 
@@
   Another valuable piece of information for understanding our patients is seeing the range of the dataset. In Figure~\ref{fig:histogram}, we
-  are able to see the spread of each predictor in our dataset. We can see that the majority of our patients are in the 20 to 30 year old range. 
-  The bloodpressure predictor provides us with a somewhat normal distribution with the average being around 70 beats per minute. The BMI
+  are able to see the spread of each predictor in our dataset. We can see that the majority of our patients are in the 20 to 30 year old range and that the distribution is skewed to the right. 
+  The bloodpressure predictor provides us with a slightly normal distribution with the average being around 70 beats per minute. The BMI
   distribution is somewhat normal as well, with a slight right skew. Whereas the diabetes pedigree function is very skewed to the right. Glucose

### 2023-12-14 code/Confusion Matrix Heatmap.R
update of confusion Matrix

@@ -9,6 +9,6 @@ ggplot(as.data.frame.table(conf_matrix), aes(x = Actual, y = Predicted, fill = F
   geom_tile() +
   scale_fill_gradient(low = "white", high = "blue") +
   geom_text(aes(label = Freq), vjust = 1) +
-  labs(title = "Confusion Matrix Heatmap", x = "Actual Outcome", y = "Predicted Outcome")
+  labs(x = "Actual Outcome", y = "Predicted Outcome")
 
 

### 2023-12-15 manuscript/manuscript.tex
update of confusion Matrix
Some reformatting
Addition of keywords and confusion matrix explanation

@@
 \usepackage[colorlinks=true, citecolor=blue]{hyperref}
+\usepackage{indentfirst}
 
@@
 \end{abstract}
+
+\section*{Keywords}
+Decision Trees, Diabetes, Machine Learning,
 
@@
 
+\section{Results}
+\label{sec:resu}
+
+After completing our methods section, we can now move on to talk about the results of our model. Based on Figure~ref{fig:importance}, we can 
+see that the top three predictors are glucose, BMI, and age. The result of glucose being of the greatest importance makes complete sense with 
+medical knowledge of diabetes as a patient that has high amounts of glucose in their bloodstream could have type 2 diabetes. This is known due
+to glucose not being absorbed by insulin as these patients do not use insulin properly. We can see that almost all of the outcomes are explained
+by glucose. The next predictor that is successful in predicting a large amount of the data is the patients BMI. This alligns on a medical standpoint
+as patients who have larger BMI values are more susceptable to getting type 2 diabetes as they are considered to be obese. Based on our results, BMI
+was able to predict about half of the outcomes. The last predictor that proved to be valuable in predicting if someone has diabetes is age. This result
+is somewhat medically backed but not entirely a result that we expected. Often times people that are older have diabetes due to numerous factors but it is 
+not as common for younger folks.
+
+\begin{figure}[tbp]
+  \centering
+  \caption{Variance Importance Plot}
+  \includegraphics[width=\textwidth]{Variance Important.pdf}
+  \label{fig:importance}
+\end{figure}
+
+Looking at Table~ref{tab:dtr}, we can see that the accuracy of our model is 0.82, this is pretty high so it will be satisfactory. This tells us that we 
+classify correctly 82 percent of the time out of the total instances. The next metric we will look at is precision, we were able to get to 72 percent.
+This is the ratio of correctly predicted positive observations to the total predicted positives. Looking at recall, we were able to correctly identify about
+74 percent of the actual positive instances. This is pretty good and will be considered satisfactory. The final metric that we have is the F1 score, this is the
+harmonic mean of precision and recall, which we were able to get a value of 0.73, this indicates a balanced performance.
+
 \begin{table}[ht]
@@
 
+Finally we can now look at Figure~ref{fig:Heatmap}, to see the confusion matrix heatmap for our decision tree analysis. Out of the 371 patients that had an outcome
+of not having diabetes, the model correctly identified 318 of them. Whereas it only incorrectly said that they did not have diabetes when in fact they did 53 times. 
+In terms of predicting that the patients had diabetes, the model predicted 136 patients had diabetes out of the 183 that actually did. In turn it precicted that 47 of the
+patients did not have diabetes when in fact they did. This shows that the model does a pretty good job in identify if the patient has diabetes but does an even better job
+identifying when the patient does not have diabetes.
+
 \begin{figure}[tbp]
@@
 \end{figure}
-
-  \begin{figure}[tbp]
-    \centering
-    \caption{Variance Importance Plot}
-    \includegraphics[width=\textwidth]{Variance Important.pdf}
-    \label{fig:importance}
-  \end{figure}
-
-\section{Results}
-\label{sec:resu}
-
-After completing our methods section, we can now move on to talk about the results of our model. Based on Figure~ref{fig:importance}, we can 
-see that the top three predictors are glucose, BMI, and age. The result of glucose being of the greatest importance makes complete sense with 
-medical knowledge of diabetes as a patient that has high amounts of glucose in their bloodstream could have type 2 diabetes. This is known due
-to glucose not being absorbed by insulin as these patients do not use insulin properly. We can see that almost all of the outcomes are explained
-by glucose. The next predictor that is successful in predicting a large amount of the data is the patients BMI. This alligns on a medical standpoint
-as patients who have larger BMI values are more susceptable to getting type 2 diabetes as they are considered to be obese. Based on our results, BMI
-was able to predict about half of the outcomes. The last predictor that proved to be valuable in predicting if someone has diabetes is age. This result
-is somewhat medically backed but not entirely a result that we expected. Often times people that are older have diabetes due to numerous factors but it is 
-not as common for younger folks.
-
-Looking at Table~ref{tab:dtr}, we can see that the accuracy of our model is 0.82, this is pretty high so it will be satisfactory. This tells us that we 
-classify correctly 82 percent of the time out of the total instances. The next metric we will look at is precision, we were able to get to 72 percent.
-This is the ratio of correctly predicted positive observations to the total predicted positives. Looking at recall, we were able to correctly identify about
-74 percent of the actual positive instances. This is pretty good and will be considered satisfactory. The final metric that we have is the F1 score, this is the
-harmonic mean of precision and recall, which we were able to get a value of 0.73, this indicates a balanced performance.
 

### 2023-12-15 manuscript/manuscript.tex
Grammatical edits throughout entire paper

@@
 \begin{abstract}
-Each year, many patients are given the diagnosis of having diabetes and are left wondering what they could have done to prevent it. This
+Each year, many patients learn that they have diabetes and are left wondering what could have done to prevent the diagnosis. This
 paper uses decision tree modeling to investigate which predictors of the Healthcare Diabetes dataset from kaggle are important in predicting 
-an outcome of diabetes. In this study, after various exploratory data analysis, the data will be split into a training and testing set
+an outcome of diabetes. In this study, after exploratory data analysis, the data was split into a training and testing set
 that will be used to build a decision tree. The performance was evaluated using standard classification metrics, including accuracy, precision, 
@@
 
-    There are two types of diabetes that the majority of patients have, type 1 or 2. The bodies of people that have type 1 diabetes
-    completely do not make insulin, so they must take a daily insulin shot. Whereas individuals who have type 2 diabetes, produce insulin but 
-    their bodies do not use it properly so they are treated either by exercise, taking on a diet or in some cases use of various medicines and/or insulin (\cite{JDC2023Difference}). 
+    There are two types of diabetes that the majority of patients have: type 1 or 2.
+The bodies of people that have type 1 diabetes do not make any insulin so they must take a daily insulin shot.
+Whereas individuals who have type 2 diabetes produce insulin, but
+    their bodies do not use it properly so they are treated either by exercise, taking on a diet, or in some cases use of various medicines and/or insulin (\cite{JDC2023Difference}). 
     Therefore, it is very important that your body has a normal glucose level. In this paper, I will be working to determine the pre-existing
     factors that can be used in order to predict if a person may have diabetes. Establishing the predictors is very important to me as many of
-    my family members have a long history of suffering from this disease. Therefore, this information could directly allow myself and my family members be
+    my family members have a long history of suffering from this disease. Therefore, this information could directly allow myself and my family members to be
     proactive in terms of prevention and care. It is important to better predict this disease to save people from unnecessary suffering.
 
-    Previously, the prediction of diabetes has been studied using other machine learning models. Specifically in Sisodia's paper, it focused
+    Previously, the prediction of diabetes has been studied using alternative machine learning models. Specifically in Sisodia's paper, it focused
     on the "Prediction of Diabetes using Classification Algorithms", they worked through the topic by producing results using support 
-    vector machine and Naive Bayes classifier. In Sisodia's paper they were able to get to an accuracy level of 76.3 percent, which we hope to exceed using
+    vector machine and Naive Bayes classifier. In Sisodia's paper they were able to get to an accuracy level of 76.3 percent, which I hope to exceed using
     a different machine learning model (\cite{Sisodia2018Prediction}). In this paper, I will be 
-    continuing the investigation with a different data set using decision trees to better understand the predictors that help to determine if a patient
-    may have diabetes.
+    continuing the investigation with a different data set using decision trees to better understand the predictors of diabetes.
 
     In this paper the specific research question that I will be focusing on is: how can we use machine learning models, specifically decision trees, in order to identify 
-    individuals that may have already or are at risk at developing diabetes? This will be researched using a data set that contains various 
+    individuals that may have already been or are at risk at developing diabetes? This will be researched using a data set that contains various 
     metrics on one's health that will be described in the next section of the paper.
@@
     In order to study the proposed research question on the topic of diabetes, I searched through many databases in order to find a set 
-    that had all of the information I was looking for. After exploration, I was able to find a data set entitled "Healthcare Diabetes Data set" on 
-    the website Kaggle. The data set was originally sourced from the National Institute of Diabetes and Digestive and Kidney Diseases. The 
-    data encompasses eight different predictors including pregnancies, glucose, blood pressure, skin thickness, insulin, BMI, diabetes 
-    pedigree function, and age. In Table~\ref{tab:head_of_data} I will provide the first couple lines of dataset in order to provide
-    more information.
+    that had all of the information I was looking for. I was searching for a data set including health metrics that included correlations to diabetes such as weight 
+    or glucose levels. This led me to further research on the impact of BMI on diabetes patients. This then became a predictor for my data set. After exploration, I was
+    able to find a data set entitled "Healthcare Diabetes Data set" on the website Kaggle. The data set was originally sourced from the National Institute of Diabetes and 
+    Digestive and Kidney Diseases. The data encompasses eight different predictors including: pregnancies, glucose, blood pressure, skin thickness, insulin, BMI, diabetes 
+    pedigree function, and age. In Table~\ref{tab:head_of_data} I will provide the first few lines of the data set in order to provide
+    more information on these predictors and their values.
 
@@
     simply be the entry's age at the time of the collection. Following this paragraph, I will provide the link to the dataset that I will 
-    be using as well as showing some descriptive statistics in Table~\ref{tab:ds} including the mean, standard deviation, and median.
+    be using as well as including some descriptive statistics as seen in Table~\ref{tab:ds} including the mean, standard deviation, and median.
 
@@
   are able to see the spread of each predictor in our dataset. We can see that the majority of our patients are in the 20 to 30 year old range and that the distribution is skewed to the right. 
-  The bloodpressure predictor provides us with a slightly normal distribution with the average being around 70 beats per minute. The BMI
-  distribution is somewhat normal as well, with a slight right skew. Whereas the diabetes pedigree function is very skewed to the right. Glucose
+  The blood pressure predictor provides us with a slightly normal distribution with the average being around 70 beats per minute. The BMI
+  distribution is somewhat normal as well, with a slight right skew. The diabetes pedigree function, however, is very skewed to the right. Glucose
   levels appear to be normal. Insulin levels appear to be very skewed to the right with the majority of values being 0. The outcome distribution
-  is exactly as expected with only values of 0 or 1, from here on values of 0 will denote no detection of diabetes and 1 will detect diabetes. The
-  number of pregnancies predictor also supplied a right skew but we know this to be valid as most people are not having 4 or more children. The 
+  is exactly as expected with only values of 0 or 1. The
+  number of pregnancies predictor also supplied a right skew but we know this to be valid as most people are not having four or more children, but there are some outliers. The 
   final predictor graphed was the skin thickness, this predictor displayed a right skew.
@@
     To begin, I will provide an explanation of what a decision tree is and why it is useful in this case. A decision tree is considered
-    to be example of supervised learning used for regression and classification. Through our analysis we will present trees, these trees
-    begin at the top with a decision root node, which is the point where you make a decision. Following down there are subsequent sub trees
-    that have other decision nodes. Finally at the bottom of each tree you reach a leaf node, where it is the output of the decision and 
+    to be an example of supervised learning used for regression and classification. Through my analysis I will present a tree, this tree
+    will begin at the top with a decision root node, which is the point where you make a decision. Following down there are subsequent sub trees
+    that have other decision nodes. Finally, at the bottom of each tree you reach a leaf node, where it is the output of the decision and 
     there are no further branches off of this.
@@
     encompass 80 percent of the data whereas the test set will only have 20 percent. This method of splitting will help to prevent any biases that may 
-    be attributed from imbalances in the distribution of critical features. After creating these two sets we will now be ready
-    to build the model, this will be done by using the rpart function in R, where method will be 'class' as this is a binary model. We will
+    be attributed from imbalances in the distribution of critical features. After creating these two sets I will now be ready
+    to build the model. This will be done by using the rpart function in R, where the method will be 'class' as this is a binary model. We will
     then see in Figure~\ref{fig:structure} the structure of the decision tree created. Looking at the leftmost path in the decision tree we 
@@
 to glucose not being absorbed by insulin as these patients do not use insulin properly. We can see that almost all of the outcomes are explained
-by glucose. The next predictor that is successful in predicting a large amount of the data is the patients BMI. This alligns on a medical standpoint
-as patients who have larger BMI values are more susceptable to getting type 2 diabetes as they are considered to be obese. Based on our results, BMI
+by glucose. The next predictor that is successful in predicting a large amount of the data is the patient's BMI. This aligns on a medical standpoint
+as patients who have larger BMI values are more susceptable to getting type 2 diabetes as they are considered to be obese. Based on my results, BMI
 was able to predict about half of the outcomes. The last predictor that proved to be valuable in predicting if someone has diabetes is age. This result
-is somewhat medically backed but not entirely a result that we expected. Often times people that are older have diabetes due to numerous factors but it is 
-not as common for younger folks.
+is somewhat medically backed but not entirely a result that I expected. Often times people that are older have diabetes due to numerous factors but it is 
+not as common for younger patients.
 
@@
 
-Looking at Table~ref{tab:dtr}, we can see that the accuracy of our model is 0.82, this is pretty high so it will be satisfactory. This tells us that we 
-classify correctly 82 percent of the time out of the total instances. The next metric we will look at is precision, we were able to get to 72 percent.
+Looking at Table~ref{tab:dtr}, we can see that the accuracy of our model is 0.82, which is high so it will be satisfactory. This tells me that we 
+classify correctly 82 percent of the time out of the total instances. The next metric we will look at is precision. We were able to get to 72 percent.
 This is the ratio of correctly predicted positive observations to the total predicted positives. Looking at recall, we were able to correctly identify about
-74 percent of the actual positive instances. This is pretty good and will be considered satisfactory. The final metric that we have is the F1 score, this is the
-harmonic mean of precision and recall, which we were able to get a value of 0.73, this indicates a balanced performance.
+74 percent of the actual positive instances. This is pretty strong and will be considered satisfactory. The final metric that we have is the F1 score, which is the
+harmonic mean of precision and recall, which we were able to get a value of 0.73. This indicates a balanced performance.
 
@@
 Finally we can now look at Figure~ref{fig:Heatmap}, to see the confusion matrix heatmap for our decision tree analysis. Out of the 371 patients that had an outcome
-of not having diabetes, the model correctly identified 318 of them. Whereas it only incorrectly said that they did not have diabetes when in fact they did 53 times. 
-In terms of predicting that the patients had diabetes, the model predicted 136 patients had diabetes out of the 183 that actually did. In turn it precicted that 47 of the
-patients did not have diabetes when in fact they did. This shows that the model does a pretty good job in identify if the patient has diabetes but does an even better job
+of not having diabetes, the model correctly identified 318 of them. The model only incorrectly missed a positive diagnosis 53 times. 
+In terms of predicting that the patients had diabetes, the model predicted 136 patients had diabetes out of the 183 that actually did. In turn it predicted that 47 of the
+patients did not have diabetes when in fact they did. This shows that the model does a pretty good job in identifying if the patient has diabetes but does an even better job
 identifying when the patient does not have diabetes.
@@
 
-Currently, this study is limited as we are set within the bounds of this data set. In future studies we can include more predictors in 
-order to see if there are any other factors that could be used to predict diabetes. As well as having an even larger data set in order to 
-ensure our accuracy. Other data sets may suggest that certain predictors are not helpful in predicting diabetes that our model did not pick 
-up on due to the metrics of our data.
+Currently, this study is limited as we are set within the bounds of this data set. In future studies I can include more predictors in 
+order to see if there are any other factors that could be used to predict diabetes. Moreover, I can have an even larger data set in order to 
+ensure my accuracy. Other data sets may suggest that certain predictors are not helpful in predicting diabetes that my model did not pick 
+up on due to the metrics of my data.
 
-This study be expanded further by using other machine learning techniques to study the data, this can be used to cross-validate our results to ensure
-that we are providing the most accurate information. Other machine learning techniques could include random forest and logistic regression, to give another
+This study can be expanded further by using other machine learning techniques to study the data. This can be used to cross-validate my results to ensure
+that I am are providing the most accurate information. Other machine learning techniques could include random forest and logistic regression to give even more
 understanding of the data.
 
-As this is a paper dealing with sensitive medical data, we have many limitations in order to be ethical as well as respecting patient 
-privacy. Based on the data set collected we should be able to ensure patient privacy pretty well as names or any very unique identifiers 
-are not listed with each entry. We simply have age for each entry so we do not have to worry as much for the sake of privacy as this cannot be traced
-back to a particular person. In terms of our direct models we may have to disregard some assumptions as they may not fit with real world data. If something unexpected 
-happens we may have to reconsider the data set used and consider finding a new one. This will be done by a case by case scenario and will 
+As this is a paper dealing with sensitive medical data, I have many limitations in order to be ethical as well as respecting patient 
+privacy. Based on the data set collected I should be able to ensure full patient privacy as names or any very unique identifiers 
+are not listed with each entry. I simply have age for each entry so I do not have to worry as much for the sake of privacy as this cannot be traced
+back to a particular person. In terms of our direct models I may have to disregard some assumptions as they may not fit with real world data. If something unexpected 
+happens, I may have to reconsider the data set used and consider finding a new one. This will be done by a case-by-case scenario and will 
 be investigated thoroughly before making any decisions.

### 2023-12-15 manuscript/manuscript.tex
Edit to decision tree descripition

@@
     to build the model. This will be done by using the rpart function in R, where the method will be 'class' as this is a binary model. We will
-    then see in Figure~\ref{fig:structure} the structure of the decision tree created. Looking at the leftmost path in the decision tree we 
-    can see that at the top the outcome was 0 meaning that the patient did not have diabetes 35 percent of the time. Moving down to the next node, 
-    we can see that 68 percent of patients with glucose level less than 128, did not have diabetes 19 percent of the time. To finish this path,
-    61 percent of the time if the patient was older than 29 years old, then 8 percent of the time the patient had diabetes. The rest of the structure
-    is similar but with different values and predictors.
+    then see in Figure~\ref{fig:structure} the structure of the decision tree created. Starting at the root node at the top of the graph, it shows the overall probability
+    of having diabetes in this dataset. 35 percent of the patients were classified as having diabetes (or the outcome of 0). The node then asks if the the
+    glucose level of the patient is less than 128. If yes, then it moves down to the left child node. 63 percent have glucose level less than 128 with a diabetes
+    outcome of 0 probability of 19 percent. The child node then asks if the patients age is less than 29 years old. If yes, the chance of not having diabetes is eight percent.
+    The rest of the tree is described similiarly but with different values of probability, predictors, and outcomes.
 

### 2023-12-15 manuscript/manuscript.tex
Minor edits to formatting

@@
 
-After completing our methods section, we can now move on to talk about the results of our model. Based on Figure~ref{fig:importance}, we can 
+After completing our methods section, we can now move on to talk about the results of our model. Based on Figure~\ref{fig:importance}, we can 
 see that the top three predictors are glucose, BMI, and age. The result of glucose being of the greatest importance makes complete sense with 
@@
 
-Looking at Table~ref{tab:dtr}, we can see that the accuracy of our model is 0.82, which is high so it will be satisfactory. This tells me that we 
+Looking at Table~\ref{tab:dtr}, we can see that the accuracy of our model is 0.82, which is high so it will be satisfactory. This tells me that we 
 classify correctly 82 percent of the time out of the total instances. The next metric we will look at is precision. We were able to get to 72 percent.
